---
title: "Coding for Random Adventures"
author: "Brian"
date: "2/25/2023"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(dataRetrieval)
library(dygraphs)
library(xts)
library(revealjs)

library(snotelr)
library(riem)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(dataRetrieval)
library(lubridate)
library(sf)
library(ggthemes)
library(xts)
library(dygraphs)
library(scales)
library(openair)
library(plotly)
library(SciViews)
knitr::opts_chunk$set(message = F, 
                      warning = F,
                      cache = T)
knitr::opts_chunk$set(echo = TRUE)
library(trend)
library(nhdplusTools)
library(lfstat)
library(ggpubr)
library(kableExtra)

#Stats
library(forcats)
library(stringr)
library(trend)
```

# Dirty Devil River

#### NWIS data

```{r downloader}
q <- readNWISdv(siteNumbers = '09333500',
                parameterCd = '00060',
                startDate = '2002-01-01',
                endDate = '2022-03-20') %>%
  rename(q = 'X_00060_00003') %>% 
  select(-X_00060_00003_cd)
```

####  Discharge, 2002-2022

```{r Static, warning = FALSE, fig.width = 8, fig.height = 5}
ggplot(q, aes(x = Date, y = q)) + 
  geom_line() + 
  ylab('Q (cfs)') + 
  ggtitle('Discharge For the Dirty Devil River, UT')
```

#### Interactive Data Plotter

```{r data plotter}
q_xts <- xts(q$q, order.by = q$Date)
dygraph(q_xts) %>%
  dyAxis("y", label = "Discharge (cfs)") 
```

#### March Discharge

```{r March clean}

q_march <- q %>% 
  addWaterYear() %>%
  group_by(waterYear)%>% 
  mutate(waterDay = (as.integer(difftime(Date, ymd(paste0(waterYear - 1 ,'-09-30')), units = "days")))) %>% filter(waterDay >= 162, waterDay <= 168) %>% 
  group_by(waterDay) %>% 
  mutate(avg = mean(q))


```

#### March Discharge, 2002-2022

```{r March, warning = FALSE, fig.width = 8, fig.height = 5}
ggplot(q_march) + 
  geom_line(mapping = aes(x = waterDay, q, color=waterYear, group=waterYear), size = 1)+ 
  geom_line(mapping = aes(x = waterDay, y = avg), size = 1.5)+
  #geom_smooth(method = lm) +
  scale_x_log10() +
  theme_few() +
  scale_colour_gradient2(low = "blue", mid = "yellow" , high = "red", 
                         midpoint=2012)+
  ylab('Q (cfs)')+
  xlab('Water Day')
```

**Discharge (cfs) for 2002-2022 for March 11-17. Average discharge is the black line.**

# Kokopelli SNOTEL stations

SNOTEL station 572 in the La Sal mountains

```{read in MOAB area SNOTELr,eval=FALSE, include=TRUE}

SNOTEL_MOAB_area <- snotel_download(site_id = c(1215, 572), path = tempdir('../data'), internal = TRUE)

write.csv(SNOTEL_MOAB_area,"C:/Users/13074/Documents/ESS580/thesis_project/the_Dirty_Devil/data/snotel_MOAB.csv", row.names = FALSE) #write in the raw data

```

```{r MOAB SNOTEL read in}
SNOTEL_MOAB_area <- read.csv("C:/Users/13074/Documents/ESS580/thesis_project/the_Dirty_Devil/data/snotel_MOAB.csv", header = TRUE)
```

```{r May weather}
SNOTEL_MOAB_area$Date <- as.Date(SNOTEL_MOAB_area$date)

str(SNOTEL_MOAB_area)                                 
                                 
                                 
SNOTEL_MOAB_area_clean <- SNOTEL_MOAB_area %>% # filter for the timeframe
  addWaterYear() %>% 
  mutate(daymonth = format(as.Date(Date), "%d-%m")) %>% 
  na.omit()


SNOTEL_MOAB_area_clean <- SNOTEL_MOAB_area_clean %>% 
  group_by(waterYear)%>% 
  mutate(waterDay = (as.integer(difftime(Date, ymd(paste0(waterYear - 1 ,'-09-30')), units = "days"))))

may_moab_572 <- SNOTEL_MOAB_area_clean %>%
  filter(site_id == 572) %>% 
  filter(waterDay >= 220 & waterDay <= 243)

#str(may_moab_572)
```

### SWE 

```{r May SWE plot}
ggplot(may_moab_572) +
  geom_line(mapping = aes(x = waterDay, y=snow_water_equivalent, color=waterYear, group=waterYear), size = 1)+ 
  scale_x_log10() +
  theme_few() +
  scale_colour_gradient2(low = "blue", mid = "yellow" , high = "red", 
                         midpoint=2011)+
  ylab('SWE (mm)')+
  xlab('Water Day')

```

*Snow water equivalent (mm) at La Sal Mountain SNOTEL (572) for May 8- May 31, 1980-2022.*

### Temperature 

```{r May temp plot}
ggplot(may_moab_572) +
  geom_line(mapping = aes(x = waterDay, y= temperature_mean, color=waterYear, group=waterYear), size = .7)+ 
  scale_x_log10() +
  theme_few() +
  scale_colour_gradient2(low = "blue", mid = "yellow" , high = "red", 
                         midpoint=2011)+
  ylab('Temperature (°C)')+
  xlab('Water Day')

```

*Daily average temperature (°C) at La Sal Mountain SNOTEL (572) for May 8- May 31, 1980-2022.*



# Fort Collins Weirdness

*Checking to see whether Fort Collins has been increasing or decreasing in daily temperature variability over the last ~42 years.*

```{r FOCO read in}

CO3005_CO3006 <- read.csv("C:/Users/13074/Downloads/CO3005_CO3006.txt", comment.char="#")

```

```{r station clean}

CO3005_CO3006$Date <- ymd(CO3005_CO3006$day)

COOP_stations_clean <- CO3005_CO3006 %>% # filter for the timeframe
  filter(station == "CO3005") %>% 
  addWaterYear() %>%
  mutate(daymonth = format(as.Date(Date), "%d-%m")) %>% 
  group_by(waterYear)%>% 
  mutate(waterDay = (as.integer(difftime(Date, ymd(paste0(waterYear - 1 ,'-09-30')), units = "days")))) %>%   na.omit()

COOP_stations_clean <- COOP_stations_clean %>% 
  mutate(avg_T_c = (highc+lowc)/2) %>% 
  filter(waterYear >= 1980 & waterYear <= 2022)


write.csv(COOP_stations_clean,"C:/Users/13074/Documents/ESS580/thesis_project/the_Dirty_Devil/data/FC_COOP_stations_clean.csv", row.names = FALSE)

ggplot(COOP_stations_clean, aes(x = Date, y = avg_T_c)) + #, color = station_name)) +
  geom_line() + #lwd = 2) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('Daily temperature (°C)') + 
  xlab('Date')
```

Looking at whether there is a increase or decrease in daily temperature variability, 1980 to 2022

```{r CO3005 detrending data}
#average water year temperature
CO3005_yearly_wy_aver <- COOP_stations_clean %>%
  filter(station == "CO3005") %>% 
  group_by(waterYear) %>%
  filter(waterYear >= 1980) %>% 
  mutate(aver_ann_temp = mean(avg_T_c))


```

```{r CO3005 detrending data2}
#Average temperature by day for all water years:
CO3005_daily_wy_aver <- CO3005_yearly_wy_aver %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(avg_T_c))
#average mean temperature by day for the period of record:
CO3005_daily_wy_aver <- CO3005_daily_wy_aver %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(CO3005_daily_wy_aver$aver_day_temp))
#str(CO3005_daily_wy_aver)
```

```{r CO3005 Figure all year average temp}
# try to show all years as means. 
CO3005_daily_wy_aver2 <- CO3005_daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  group_by(waterDay) %>%
  mutate(date_temp = mean(avg_T_c))
  
CO3005_daily_wy_aver2$date_temp <- signif(CO3005_daily_wy_aver2$date_temp,3) #reduce the sig figs
ggplot(CO3005_daily_wy_aver2, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  ylab('Average Daily temperature (°C)') + 
  xlab('Day of water year')
```

Standard deviation- Determining residuals

```{r CO3005 residuals}
CO3005_standard_dev <- CO3005_daily_wy_aver %>% 
  group_by(waterYear) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+avg_T_c-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))
#checking the detrend
detrend <- CO3005_standard_dev %>% 
  filter(waterYear == 1984)

ggplot(detrend, aes(x = waterDay, y = residual))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  geom_hline(yintercept=0, color = "red", size = 1) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  ylab('Average Daily temperature (°C)') + 
  xlab('Day of water year')
mean(CO3005_standard_dev$residual)
```
*The mean of the residuals is close enough to zero*

Calculating standard deviation for the timeseries
```{r 444 CO3005 sd update}
CO3005_standard_dev_all <- CO3005_standard_dev %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())
CO3005_standard_dev_all <- CO3005_standard_dev_all %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)
CO3005_standard_dev_all %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')
```


```{r 444 CO3005 std dev update}
#CALLING THIS something different
CO3005_all_V2 <- ggplot(CO3005_standard_dev_all, aes(x = waterYear, y = sd_2))+
  geom_line(size= 0.7) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')
print(CO3005_all_V2)
```


```{r  CO3005 sd }
CO3005_sd_mk <- mk.test(CO3005_standard_dev_all$sd_2)
print(CO3005_sd_mk)
CO3005_sd_sens <- sens.slope(CO3005_standard_dev_all$sd_2)
print(CO3005_sd_sens)
```

Summer temperature standard deviation

```{r CO3005 summer}
CO3005_standard_dev_all_summer <- CO3005_standard_dev %>%
  filter(waterDay >= 244 & waterDay <= 335) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(nmbr = n())
CO3005_standard_dev_all_summer <- CO3005_standard_dev_all_summer %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)
CO3005_standard_dev_all_summer %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')
ggplot(CO3005_standard_dev_all_summer, aes(x = waterYear, y = sd_2))+
  geom_line(size= 0.7) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')
```

**CO3005 Jun-Aug standard deviation for water years 1984-2022**

Mann-Kendall & Sen’s Slope

Summer standard deviations.
```{r CO3005 sd mk & ss summer}
CO3005_sd_mk_summer <- mk.test(CO3005_standard_dev_all_summer$sd_2)
print(CO3005_sd_mk_summer)
CO3005_sd_sens_summer <- sens.slope(CO3005_standard_dev_all_summer$sd_2)
print(CO3005_sd_sens_summer)
```

Winter temperature standard deviation

```{r CO3005 winter}
CO3005_standard_dev_all_winter <- CO3005_standard_dev %>%
  filter(waterDay >= 32 & waterDay <= 182) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(nmbr = n())

CO3005_standard_dev_all_winter <- CO3005_standard_dev_all_winter %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)
CO3005_standard_dev_all_winter %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')
ggplot(CO3005_standard_dev_all_winter, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')
```

**CO3005 Nov-Mar standard deviation for water years 1984-2022**

Mann-Kendall & Sen’s Slope

Winter standard deviations.
```{r CO3005 sd mk & ss winter}
CO3005_sd_mk_winter <- mk.test(CO3005_standard_dev_all_winter$sd_2)
print(CO3005_sd_mk_winter)
CO3005_sd_sens_winter <- sens.slope(CO3005_standard_dev_all_winter$sd_2)
print(CO3005_sd_sens_winter)
```


Spring and Fall temperature standard deviation


```{r CO3005 spring}
CO3005_standard_dev_all_spring <- CO3005_standard_dev %>%
  filter(waterDay >= 183 & waterDay <= 243) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(nmbr = n())
CO3005_standard_dev_all_spring <- CO3005_standard_dev_all_spring %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)
CO3005_standard_dev_all_spring %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')
ggplot(CO3005_standard_dev_all_spring, aes(x = waterYear, y = sd_2))+
  geom_line(size= 0.7) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')
```

**CO3005 Jun-Aug standard deviation for water years 1984-2022**

Mann-Kendall & Sen’s Slope

Spring standard deviations.
```{r CO3005 sd mk & ss spring}
CO3005_sd_mk_spring <- mk.test(CO3005_standard_dev_all_spring$sd_2)
print(CO3005_sd_mk_spring)
CO3005_sd_sens_spring <- sens.slope(CO3005_standard_dev_all_spring$sd_2)
print(CO3005_sd_sens_spring)
```

Fall temperature standard deviation

```{r CO3005 fall}
CO3005_standard_dev_all_fall <- CO3005_standard_dev %>%
  filter(waterDay >= 336 | waterDay <= 31) %>% # this might be better off as daymonth rather than day of water year due to leap year
  group_by(waterYear) %>% 
  mutate(nmbr = n())
# Nope. This did some weird stuff with twice the observations.
# CO3005_standard_dev_all_fall <- CO3005_standard_dev %>%
#   filter(daymonth >= "01-11" & daymonth <= "31-03") %>%
#   group_by(waterYear) %>% 
#   mutate(nmbr = n())
CO3005_standard_dev_all_fall <- CO3005_standard_dev_all_fall %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)
CO3005_standard_dev_all_fall %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')
ggplot(CO3005_standard_dev_all_fall, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')
```

**CO3005 Nov-Mar standard deviation for water years 1984-2022**

Mann-Kendall & Sen’s Slope

Fall standard deviations.
```{r CO3005 sd mk & ss fall}
CO3005_sd_mk_fall <- mk.test(CO3005_standard_dev_all_fall$sd_2)
print(CO3005_sd_mk_fall)
CO3005_sd_sens_fall <- sens.slope(CO3005_standard_dev_all_fall$sd_2)
print(CO3005_sd_sens_fall)
```


